---
title: "Untitled"
author: "Andrew Wells"
date: "1/13/2023"
output: html_document
---
The code R code below was performed in order to analyze data from Malawi's public infrastructure procurement portal (IPPI). The code mainly demonstrates my ability to clean data with R. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("tidyverse")
#install.packages("readr")

library(tidyverse)
library(readr)
```

```{r}
library(httr)
library(jsonlite)

#Extracting IPPI data through the API
api_url <- "https://ippi.mw/api/projects/query"
response <- POST(api_url, body = list(start_date = "2021-01-01", end_date = "2023-01-16"), 
                 encode = "json")
d <- fromJSON(content(response, as="text"))
df <- as.data.frame(d$projects)
write.csv(df, file = "ippi_2023_01_16.csv", 
          row.names = FALSE, fileEncoding = "UTF-8")
```


```{r}
#Data must be loaded manually
ippi_2023_01_16 <- read_csv("~/Desktop/Jobs/Malawi/Project Implementation/Initiatives:project components/IPPI/ippi_report/data/ippi_2023_01_16.csv")

#Resetting df name
df_raw <- ippi_2023_01_16
```

```{r}
#Initial inspection of information completed

df_raw <- df_raw %>%
  select(-c(...1))

df_raw$completion_count <- apply(df_raw, 1, function(x) sum(!is.na(x)))  

df_raw %>%
  filter(completion_count == 14)

mean(df_raw$completion_count)

sapply(df_raw, function(x) sum(!is.na(x)))
```


```{r}
#Column that counts the number of NAs in each row

df$na_count <- apply(df, 1, function(x) sum(is.na(x)))
```

#Cleaning data

##Cleaning/creating dates columns
```{r}
#Setting variables as dates
df$period.startDate <- as.Date(df$period.startDate)
df$period.endDate <- as.Date(df$period.endDate)
df$period.maxExtentDate <- as.Date(df$period.maxExtentDate)

#Creating variables for planned project duration and extension period
df <- df %>%
  mutate(planned_proj_days = difftime(period.endDate, period.startDate, units = "days")) %>%
  mutate(max_extend_days = difftime(period.maxExtentDate, period.endDate, units = "days"))
  
#Setting planning project days to numeric
df$planned_proj_days <- as.numeric(df$planned_proj_days)

#Setting extension period as numeric
df$max_extend_days <- as.numeric(df$max_extend_days)
df["max_extend_days"][is.na(df["max_extend_days"])] <- 0
df$max_extend_days <- as.numeric(df$max_extend_days)

#Creating variable for total project period (planned + extension)
df <- df %>%
  mutate(total_proj_days = planned_proj_days + max_extend_days)

#Setting budget date variables as dates
df$budget.requestDate <- as.Date(df$budget.requestDate)
df$budget.approvalDate <- as.Date(df$budget.approvalDate)

#Creating a budget approval timeline variable
df <- df %>%
  mutate(budget_approval_days = difftime(budget.approvalDate, budget.requestDate, units = "days"))

#Setting variables as numeric
df$budget_approval_days <- as.numeric(df$budget_approval_days)
df$total_proj_days <- as.numeric(df$total_proj_days)
```

##Cleaning sectors column
```{r}
#Removing punctuation
df$sector <- gsub("\\[|\\]", "", df$sector)
df$sector <- gsub("'","", df$sector)

#Changing variable names
df$sector <- str_replace(df$sector, "transport.road", "transportRoad")
df$sector <- str_replace(df$sector, "transport.urban", "transportUrban")

#Separating sector
df <- df %>% 
  separate(sector, c("sector_1", "sector_2", "sector_3", "sector_4", "sector_5"))

#Creating columns based on sectors
df <- df %>%
  mutate(agriculture = ifelse(sector_1 == "agriculture" | sector_2 == "agriculture" | sector_3 == "agriculture" |
                                 sector_4 == "agriculture" | sector_5 == "agriculture", 1, 0)) %>%
  mutate(communications = ifelse(sector_1 == "communications" | sector_2 == "communications" | sector_3 == "communications" |
                                   sector_4 == "communications" | sector_5 == "communications", 1, 0)) %>%
  mutate(cultureSportsAndRecreation = ifelse(sector_1 == "cultureSportsAndRecreation" | sector_2 == "cultureSportsAndRecreation" | sector_3 == "cultureSportsAndRecreation" |
                                   sector_4 == "cultureSportsAndRecreation" | sector_5 == "cultureSportsAndRecreation", 1, 0)) %>%
  mutate(economy = ifelse(sector_1 == "economy" | sector_2 == "economy" | sector_3 == "economy" |
                                   sector_4 == "economy" | sector_5 == "economy", 1, 0)) %>%
  mutate(education = ifelse(sector_1 == "education" | sector_2 == "education" | sector_3 == "education" |
                                   sector_4 == "education" | sector_5 == "education", 1, 0)) %>%
  mutate(energy = ifelse(sector_1 == "energy" | sector_2 == "energy" | sector_3 == "energy" |
                                   sector_4 == "energy" | sector_5 == "energy", 1, 0)) %>%
  mutate(governance = ifelse(sector_1 == "governance" | sector_2 == "governance" | sector_3 == "governance" |
                                   sector_4 == "governance" | sector_5 == "governance", 1, 0)) %>%
  mutate(health = ifelse(sector_1 == "health" | sector_2 == "health" | sector_3 == "health" |
                                   sector_4 == "health" | sector_5 == "health", 1, 0)) %>%
  mutate(socialHousing = ifelse(sector_1 == "socialHousing" | sector_2 == "socialHousing" | sector_3 == "socialHousing" |
                                   sector_4 == "socialHousing" | sector_5 == "socialHousing", 1, 0)) %>%
  mutate(transport = ifelse(sector_1 == "transport" | sector_2 == "transport" | sector_3 == "transport" |
                                   sector_4 == "transport" | sector_5 == "transport", 1, 0)) %>%
  mutate(transportRoad = ifelse(sector_1 == "transportRoad" | sector_2 == "transportRoad" | sector_3 == "transportRoad" |
                                   sector_4 == "transportRoad" | sector_5 == "transportRoad", 1, 0)) %>%
  mutate(transportUrban = ifelse(sector_1 == "transportUrban" | sector_2 == "transportUrban" | sector_3 == "transportUrban" |
                                   sector_4 == "transportUrban" | sector_5 == "transportUrban", 1, 0)) %>%
  mutate(waterAndWaste = ifelse(sector_1 == "waterAndWaste" | sector_2 == "waterAndWaste" | sector_3 == "waterAndWaste" |
                                   sector_4 == "waterAndWaste" | sector_5 == "waterAndWaste", 1, 0))
```
  
##Cleaning/creating budget columns
```{r}
#Creating common currency column

df$budget.amount.amount <- format(df$budget.amount.amount, scientific = FALSE)
df$budget.amount.amount <- as.numeric(df$budget.amount.amount)

df$amount_mwk <- ifelse(df$budget.amount.currency == "MWK", df$budget.amount.amount, df$budget.amount.amount * 1000)
df$amount_mwk <- format(df$amount_mwk, scientific = FALSE)
df$amount_mwk <- as.numeric(df$amount_mwk)

df$amount_usd_eur <- df$amount_mwk / 1000

#Creating over budget column
df$budget.amount.amount <- as.numeric(df$budget.amount.amount)
df$completion.finalValue.amount <- as.numeric(df$completion.finalValue.amount)

df$budget.amount.amount <- round(df$budget.amount.amount, digits = 0)
df$completion.finalValue.amount <- round(df$completion.finalValue.amount, digits = 0)

df <- df %>%
  mutate(over_budget = completion.finalValue.amount - budget.amount.amount)
```

##Cleaning/creating regions column
```{r}

df_locations <- df %>%
  select(locations)

#Cleaning data
d <- gsub(".*'region':","", df_locations$locations)

f <- as.data.frame(d)

g <- gsub('[[:punct:] ]+',' ', f$d)

h <- as.data.frame(g)

h$region <- gsub("([A-Za-z]+).*", "\\1", h$g)

#Binding with main dataframe
df <- cbind(df, h)
```

##Cleaning/creating coordinates column
```{r}
#Creating coordinates column

d <- gsub(".*'coordinates':","", df_locations$locations)

f <- as.data.frame(d)

g <- sub(" 'address'.*", "", f$d) 

h <- as.data.frame(g)

h$coordinates <- ifelse(grepl("id", h$g), NA, h$g)

h <- h %>%
  select(coordinates)

h$coordinates <- gsub("},","",h$coordinates)

#Remove all coordinates after first listing
h$coords_2 <- sub("^([^,]*,[^,]*),.*", "\\1", h$coordinates)

#Remove brackets
h$coords_2 <- gsub("\\[|\\]", "", h$coords_2)

#Extract values from coordinates variable to make lat and long variables

h$long <- sub(",.*", "", h$coords_2)
h$lat <- sub(".*, ", "", h$coords_2) 

#Joining coordinates df with full df
df <- cbind(df, h)
```

##Final cleaning of df
```{r}
df <- df %>%
  select(-c(g))
```

```{r}
write.csv(df, "/Users/andrewwells/Desktop/Jobs/Malawi/Project Implementation/Initiatives:project components/IPPI/ippi_report/data/ippi.csv")
```

##Cleaning 'party' column data, which is stored in dataset as a dictionary.

```{r}
#Creating party column

#Examining parties variable
df$parties

#Creating dataframe with just parties
df_parties <- df %>%
  select(parties)

#Splitting df so each row is a different party
a <- str_split(df_parties, "'id':")
b <- as.data.frame(a)

#Renaming column
colnames(b) <- c("party") 

#Deleting all text between "address" and "role"
c <- gsub("(?<=address).*(?=role)", " ", b$party, perl = TRUE)
d <- as.data.frame(c)
d

#Deleting all quotation marks
d$c <- gsub("'","",d$c)

#Deleting all text before the party name
d$c <- gsub(".*name:", "", d$c)

#Deleting certain unimportant words
d$c <- str_remove_all(d$c, "address")
d$c <- str_remove_all(d$c, "role")
d$c <- str_remove_all(d$c, "s:")

#Creating new variable with just the function of each party
d$e <- sub("^.*?,", "", d$c)

#Deleting all commas
d$c <- gsub(",.*","",d$c)

#Deleting all punctuation
d$f <- gsub('[[:punct:]]+','', d$e)

#Making different columns for the different roles of each party
g <- d %>% 
  separate(f, c("role", "role_1", "role_2", "role_3", "role_4", "role_5")) %>%
  select(-c("role", "e"))

#Deleting rows with NA for role_1 (no info)
g <- g %>% 
  drop_na(role_1)

#Making additional rows to cbind
pat <- paste0("\\b", "id", "\\b", collapse = "|")    

m <- df %>%
  mutate(proj2 = stringr::str_extract_all(parties, pat)) %>% 
  tidyr::unnest(proj2) %>%
  drop_na(parties)

m <- m %>%
  mutate(proj_num = ...1 + 1)
  
g <- g %>%  
  slice(-1)

#Rename "c" column to "party
g <- g %>%
  rename(party = "c")
```

```{r}
#Creating columns for each role

g <- g %>%
  mutate(administrativeEntity = ifelse(role_1 == "administrativeEntity" | role_2 == "administrativeEntity" | role_3 == "administrativeEntity" |
                                 role_4 == "administrativeEntity" | role_5 == "administrativeEntity", 1, 0)) %>%
  mutate(buyer = ifelse(role_1 == "buyer" | role_2 == "buyer" | role_3 == "buyer" |
                                   role_4 == "buyer" | role_5 == "buyer", 1, 0)) %>%
  mutate(email = ifelse(role_1 == "email" | role_2 == "email" | role_3 == "email" |
                                   role_4 == "email" | role_5 == "email", 1, 0)) %>%
  mutate(funder = ifelse(role_1 == "funder" | role_2 == "funder" | role_3 == "funder" |
                                   role_4 == "funder" | role_5 == "funder", 1, 0)) %>%
  mutate(interestedParty = ifelse(role_1 == "interestedParty" | role_2 == "interestedParty" | role_3 == "interestedParty" |
                                   role_4 == "interestedParty" | role_5 == "interestedParty", 1, 0)) %>%
  mutate(payee = ifelse(role_1 == "payee" | role_2 == "payee" | role_3 == "payee" |
                                   role_4 == "payee" | role_5 == "payee", 1, 0)) %>%
  mutate(payer = ifelse(role_1 == "payer" | role_2 == "payer" | role_3 == "payer" |
                                   role_4 == "payer" | role_5 == "payer", 1, 0)) %>%
  mutate(procuringEntity = ifelse(role_1 == "procuringEntity" | role_2 == "procuringEntity" | role_3 == "procuringEntity" |
                                   role_4 == "procuringEntity" | role_5 == "procuringEntity", 1, 0)) %>%
  mutate(supplier = ifelse(role_1 == "supplier" | role_2 == "supplier" | role_3 == "supplier" |
                                   role_4 == "supplier" | role_5 == "supplier", 1, 0)) %>%
  mutate(tenderer = ifelse(role_1 == "tenderer" | role_2 == "tenderer" | role_3 == "tenderer" |
                                   role_4 == "tenderer" | role_5 == "tenderer", 1, 0))
```

```{r}
#Creating columns for party names in roles

g <- g %>%
  mutate(administrativeEntity_name = ifelse(role_1 == "administrativeEntity" | role_2 == "administrativeEntity" | role_3 == "administrativeEntity" |
                                 role_4 == "administrativeEntity" | role_5 == "administrativeEntity", g$party, 0)) %>%
  mutate(buyer_name = ifelse(role_1 == "buyer" | role_2 == "buyer" | role_3 == "buyer" |
                                   role_4 == "buyer" | role_5 == "buyer", g$party, 0)) %>%
  mutate(email_name = ifelse(role_1 == "email" | role_2 == "email" | role_3 == "email" |
                                   role_4 == "email" | role_5 == "email", g$party, 0)) %>%
  mutate(funder_name = ifelse(role_1 == "funder" | role_2 == "funder" | role_3 == "funder" |
                                   role_4 == "funder" | role_5 == "funder", g$party, 0)) %>%
  mutate(interestedParty_name = ifelse(role_1 == "interestedParty" | role_2 == "interestedParty" | role_3 == "interestedParty" |
                                   role_4 == "interestedParty" | role_5 == "interestedParty", g$party, 0)) %>%
  mutate(payee_name = ifelse(role_1 == "payee" | role_2 == "payee" | role_3 == "payee" |
                                   role_4 == "payee" | role_5 == "payee", g$party, 0)) %>%
  mutate(payer_name = ifelse(role_1 == "payer" | role_2 == "payer" | role_3 == "payer" |
                                   role_4 == "payer" | role_5 == "payer", g$party, 0)) %>%
  mutate(procuringEntity_name = ifelse(role_1 == "procuringEntity" | role_2 == "procuringEntity" | role_3 == "procuringEntity" |
                                   role_4 == "procuringEntity" | role_5 == "procuringEntity", g$party, 0)) %>%
  mutate(supplier_name = ifelse(role_1 == "supplier" | role_2 == "supplier" | role_3 == "supplier" |
                                   role_4 == "supplier" | role_5 == "supplier", g$party, 0)) %>%
  mutate(tenderer_name = ifelse(role_1 == "tenderer" | role_2 == "tenderer" | role_3 == "tenderer" |
                                   role_4 == "tenderer" | role_5 == "tenderer", g$party, 0))
```

```{r}
#Combining columns
party <- cbind(m, g)
```

```{r}
#Check if any party is working twice on same project
party %>%
  group_by(title) %>%
  distinct(party, .keep_all=TRUE)
```

```{r}
#Save cleaned data
write.csv(party, "/Users/andrewwells/Desktop/Jobs/Malawi/Project Implementation/Initiatives:project components/IPPI/ippi_report/data/party_df.csv")
```

